DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-02-02 paper/comparison.tex
comparison section revision.

Finally, we compare the effect of co-scheduling memory bandwidth
intensive synthetic benchmarks on the DNN control loop timing. 
Figure~\ref{fig:sys_bench} shows the results.
As discussed in 
Section~\ref{sec:eval-memhog}, we observed dramatic execution time
increases, up to 11.5X, in Raspberry Pi 3 as we increased the number of
co-scheduled tasks. We also observe increased control loop execution
timing in the Intel Up and Jetson TX2, but the degree of the increase is 
not as dramatic as the Pi 3. Compared to their respective solo timings 
(i.e., the model runs on a single core in isolation), Intel UP suffers up to
2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
2.2X and 2.1X increases, respectively. This is somewhat suprising
because the Raspberry Pi 3's cores are in-order architecture based while
the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
based, and that the memory intensive tasks on out-of-order cores can
generate more memory traffic. We suspect that this is because the
memory subsystems in the Intel UP and TX2 platforms provide higher
bandwidth and fairness than the memory subsystem of the Pi 3. However, 
the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
percentage of L2 miss rates increases are more 
memory intensive co-runners are introduced. This was not the case for 
the UP board, where the model running on two cores generated the highest
percentage of L2 cache misses. This is most likely due to the CPU 
architecture of the UP's CPU which is Intel based, whereas both the Pi 3
and TX2 have ARM based CPUs.

Another interesting observation is that the TX2 (GPU) also suffers
considerable execution time increase (2.1X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU. In other words,
the DNN model has dedicated access to the GPU. This is, however, a
known characteristic of integrated CPU-GPU architectue based
platforms where both CPU and GPU share the same memory
subsystem~\cite{Ali2017}. As a result, the TX2 (GPU) fails to meet the
10ms deadline for 100 Hz control that would have been feasible if
there was no contention between the CPU cores and the GPU.

>>>>

Finally, we compare the effect of co-scheduling memory bandwidth
intensive synthetic benchmarks on the DNN control loop timing.

Figure~\ref{fig:sys_bench} shows the results. As discussed in 
Section~\ref{sec:eval-memhog}, we observed dramatic execution time
increases, up to 11.5X, in Raspberry Pi 3 as we increased the number of
co-scheduled tasks. We also observe increased control loop execution
timing in the Intel Up and Jetson TX2, but the degree of the increase is 
not as dramatic as the Pi 3. Compared to their respective solo timings 
(i.e., the model runs on a single core in isolation), Intel UP suffers up to
2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
2.2X and 2.1X increases, respectively. This is somewhat surprising
because the Raspberry Pi 3's cores are in-order architecture based while
the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
based, and that the memory intensive tasks on out-of-order cores can
generate more memory traffic. We believe that this is because the
memory subsystems in the Intel UP and TX2 platforms provide higher
performance than the memory subsystem of the Pi 3 as suggested from
the measured memory bandwidth results in Table~\ref{tbl:platforms}
('Measured Bandwidth').

Another interesting observation is that the TX2 (GPU) also suffers
considerable execution time increase (2.1X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU. In other words,
the DNN model has dedicated access to the GPU. This is, however, a
known characteristic of integrated CPU-GPU architecture based
platforms where both CPU and GPU share the same memory
subsystem~\cite{Ali2017}. As a result, the TX2 (GPU) fails to meet the
10ms deadline for 100 Hz control that would have been feasible if
there was no contention between the CPU cores and the GPU.

Figure~\ref{fig:sys_l2miss} shows the L2 miss rates of DNN control
loop. The key takeaway of this result is that the L2 miss rate is a
poor indicator to the execution time of the control loop. As shown
already, Raspberry Pi 3's increased L2 misses do not fully explain
increase in processing times. In case of Intel UP, as the number of
interfering Bandwidth benchmark instances increases, the L2 miss rates
even drop slightly. In case of TX2 (CPU), from solo to one co-runner,
L2 miss rate increase almost twice, but then the execution time
increase is only about less than 10\%. However, when there are
co-runners, even though the L2 miss rates are largely unchanged, the
DNN processing time increases significantly. TX2 (GPU) also shows
similar behavior. All of these results essentially point to the fact
that DNN inferencing workload is largely memory performance sensitive
but not L2 cache space sensitive.

%% However, the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
%% pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
%% percentage of L2 miss rates increases are more 
%% memory intensive co-runners are introduced. This was not the case for 
%% the UP board, where the model running on two cores generated the highest
%% percentage of L2 cache misses. This is most likely due to the CPU 
%% architecture of the UP's CPU which is Intel based, whereas both the Pi 3
%% and TX2 have ARM based CPUs.

### 2018-02-02 paper/intro.tex
comparison section revision.

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's CNN-based
control loop is feasible on today's embedded computing platforms, even 
as inexpensive as the Raspberry Pi 3. We find that the control loop 
completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; or we can achieve 20 Hz control frequency using
just one core of the Raspberry Pi 3's quad-core CPU.
Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

>>>>

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's CNN is
feasible on today's embedded computing platforms, even  
as inexpensive as the Raspberry Pi 3. We find that the control loop 
completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; or we can achieve 20 Hz control frequency using
just one core.
Other tested embedded platforms, the Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

### 2018-02-02 paper/overview.tex
comparison section revision.

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on tracks we created (Figure~\ref{fig:tracks}) to record
timestamped videos and contol commands. The stored data is then copied 
to a desktop computer, which is equipped with a NVIDIA GTX 1050ti GPU, 
where we train the network to accelerate training speed. 
For comparison, training the network on the Raspbeery Pi 3 takes
approximately 4 hours, whereas it takes only about 4 minutes on the
desktop computer using the GTX 1050ti GPU.

@@

Note that although the steering angle output of the network $angle$ is
a continuous real value, the RC car platform we used unfortunately
only supports three discrete angles---left (-30$^{\circ}$), center
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
Currently, we approximate the network generated real-valued
angle to the closest one of the three angles, which may
introduce inaccuracy in control.
In the future, we plan to use a different (more expensive) RC car
platform that can precisely control the car's steering angle.

>>>>

% data collection and training.
To collect the training data, a human pilot manually drives the RC car
on tracks we created (Figure~\ref{fig:tracks}) to record
timestamped videos and contol commands. The stored data is then copied 
to a desktop computer, which is equipped with a NVIDIA GTX 1050 Ti GPU, 
where we train the network to accelerate training speed. 
For comparison, training the network on the Raspbeery Pi 3 takes
approximately 4 hours, whereas it takes only about 4 minutes on the
desktop computer using the GTX 1050 Ti GPU.

@@

Note that although the steering angle output of the network ($angle$) is
a continuous real value, the RC car platform we used unfortunately
only supports three discrete angles---left (-30$^{\circ}$), center
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
Currently, we approximate the network generated real-valued
angle to the closest one of the three angles, which may
introduce inaccuracy in control.
In the future, we plan to use a different (more expensive) RC car
platform that can precisely control the car's steering angle.
